Galois Transformers and Modular Abstract Interpreters:\\ Reusable Metatheory for Program Analysis

### 2014-11-13 writeup/pldi15.tex
todos and edits
build was broken; fixed now
Updated abstract.

\newcommand\litM{M}
\newcommand\litTime{Time}
\newcommand\litStore{Store}
\newcommand\litVal{Val}

@@

The design and implementation of static analyzers have becoming
increasingly systematic.  In fact, design and implementation have
remained seemingly on the verge of full mechanization for several
years.  A stumbling block in full mechanization has been the ad hoc
nature of soundness proofs accompanying each analyzer.  While design
and implementation is largely systematic, soundness proofs can change
significantly with (apparently) minor changes to the semantics and
analyzers themselves.

\par We solve the problem of systematically constructing static
analyzers by introducing the notion of a \emph{Galois transformer}, a
monad transformer that satisfies the properties of a Galois
connection.  In concert with a monadic interpreter, we are able to
define a library of monad transfomers implementing classic context,
path, and heap (in-)senstive building blocks which can be composed
together to realize a program analysis monad, \emph{independent of the
  language being analyzed}.

\par Significantly, a Galois transformer can be proved sound once and
for all, making it a reusable analysis component.  As new analaysis
features and abstractions are developed and mixed in, soundness proofs
need not be reconstructed, as the composition of a monad transfomer
stack is sound by virtue of its consituents.  We believe Galois
transformers provide a viable foundation for resuable and composable 
metatheory for program analysis.
\end{abstract}
% }-}

>>>>

\newcommand\litM{M}
\newcommand\litTime{Time}
\newcommand\litStore{Store}
\newcommand\litVal{Val}
\newcommand\litKAddr{KAddr}
\newcommand\litKStore{KStore}
\newcommand\litEnv{Env}
\newcommand\litClo{Clo}

@@

The design and implementation of static analyzers have becoming
increasingly systematic.  In fact, for large classes of analyzers, design and
implementation have remained seemingly (and now stubbornly) on the verge of full mechanization
for several years.  A stumbling block in full mechanization has been the
\emph{ad hoc} nature of soundness proofs accompanying each analyzer.
While design and implementation is largely systematic, soundness proofs
can change significantly with what seem like minor changes to the
semantics and analyzers themselves.  An achievement of this work is to systematize, parmaterize and modularize the proofs of soundness, so as to make them composoble across analytic properties.

\par We solve the problem of systematically constructing static
analyzers by introducing the notion of a \emph{Galois transformer}, a
monad transformer that satisfies the properties of a Galois
connection.  In concert with a monadic interpreter, we are able to
define a library of monad transformers that implement and represent classic context-,
path-, and heap-(in-?)sensitive building blocks.  Moreover, these can be composed
together to realize a program analysis monad, \emph{independent of the
  language being analyzed}.

\par Significantly, a Galois transformer can be proved sound once and
for all, making it a reusable analysis component.  As new analysis
features and abstractions are developed and mixed in, soundness proofs
need not be reconstructed, as the composition of a monad transfomer
stack is sound by virtue of its constituents.  Galois
transformers provide a viable foundation for resuable and composable 
metatheory for program analysis.

Finally, these Galois transformers shift the level of abstraction in analysis design and implementation
to a level where non-specialists have the ability to synthesize sound analyzers
over a number of parameters.
\end{abstract}
% }-}
